---
title: "Federated Learning with shrinkr"
author: "Jacob M. Maronge"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Federated Learning with shrinkr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  warning = FALSE,
  message = FALSE
)
```

## Introduction

**Federated learning** enables collaborative analysis across multiple sites without centralizing data. This is critical when:

- Data cannot leave institutional firewalls (HIPAA, GDPR)
- Each site has proprietary or sensitive data
- Centralizing data is logistically infeasible
- You want to preserve patient privacy

**shrinkr's two-stage architecture naturally enables federated learning:**

```
Site 1: Stage 1 model → Posterior samples (or summaries)
Site 2: Stage 1 model → Posterior samples (or summaries)  } → Central
Site 3: Stage 1 model → Posterior samples (or summaries)  }   Coordinator
   ...                                                     }   applies
Site K: Stage 1 model → Posterior samples (or summaries)  }   Stage 2 shrinkage
```

**Data never leaves the sites**, only statistical summaries are shared.

> **Important: CLT Assumption for Summary Statistics**
>
> If sites share only summary statistics (means + SEs) rather than full posteriors, the analysis relies on the **Bayesian Central Limit Theorem**. This assumes posteriors are approximately normal, which requires:
>
> - Adequate sample sizes at each site
> - Parameters in the interior (not near boundaries)
> - Well-behaved likelihood functions
> - Regular posterior geometry
>
> **Always verify** posterior normality before using summary statistics! When in doubt, share full posteriors or send mixture approximations (e.g. send `fit_mixture()` output).

```{r packages}
library(shrinkr)
library(distributional)
library(posterior)
library(dplyr)
library(ggplot2)
library(tidyr)
```

## Use Case: Multi-Hospital Mortality Prediction

We'll analyze a federated clinical prediction model across 6 hospitals. Each hospital:   
- Has developed a logistic regression model for 30-day mortality   
- Cannot share patient-level data (HIPAA compliance)   
- Wants to improve predictions by borrowing strength across sites   

**Goal**: Combine site-specific models while respecting data governance constraints.

## Scenario Setup

### The Federated Network

```{r network_setup}
hospitals <- data.frame(
  site_id = 1:6,
  name = c("Metro General", "County Regional", "University Medical", 
           "Community Hospital", "Veterans Affairs", "Children's Specialty"),
  location = c("Urban", "Suburban", "Academic", "Rural", "Urban", "Urban"),
  n_patients = c(1500, 800, 2200, 350, 1100, 900),
  baseline_risk = c(0.15, 0.12, 0.18, 0.10, 0.20, 0.14)
)

print(hospitals)
```

### The Clinical Model

Each site fits:
```
logit(mortality) = β₀ + β₁(age) + β₂(severity_score) + β₃(comorbidities)
```

**Parameter of interest**: `β₁` (age effect on mortality)   
- We want site-specific estimates   
- But want to borrow strength across the network   

## Federated Workflow

### Step 1: Each Site Fits Independently

In practice, this happens behind each site's firewall. We simulate:

```{r site_models}
set.seed(1104)

# True network-level parameters (unknown in practice)
true_mu_age <- 0.05    # log-OR per year
true_tau_age <- 0.015  # between-site heterogeneity
true_site_effects <- rnorm(6, true_mu_age, true_tau_age)

# Simulate Stage 1: Each site fits their model independently
# In reality: glm(), stan_glm(), or other Bayesian logistic regression

site_posteriors <- list()
site_sample_sizes <- hospitals$n_patients

for(i in 1:6) {
  # Posterior for age coefficient (β₁)
  # SE inversely proportional to sqrt(sample size)
  se_i <- 0.02 * sqrt(800 / site_sample_sizes[i])
  
  site_posteriors[[hospitals$name[i]]] <- matrix(
    rnorm(4000, true_site_effects[i], se_i),
    ncol = 1
  )
}

# Each site computes summaries
site_summaries <- data.frame(
  site = hospitals$name,
  n_patients = site_sample_sizes,
  beta_age_mean = sapply(site_posteriors, mean),
  beta_age_se = sapply(site_posteriors, sd)
) %>%
  mutate(
    ci_lower = beta_age_mean - 1.96 * beta_age_se,
    ci_upper = beta_age_mean + 1.96 * beta_age_se
  )

print(site_summaries)
```

**Key observations:**   
- Smaller sites (e.g., Community Hospital) have wider intervals   
- Point estimates vary across sites (0.039 to 0.063)   
- Rural site with fewer patients is least precise   

### Step 2: Sites Share Summaries with Coordinator

Two federated learning paths are possible:

#### Path A: Share Full Posteriors (if permitted)
```{r share_posteriors}
# Sites share posterior samples (4000 draws each)
# This is more informative but requires more data transfer

cat("Data shared per site:\n")
cat("  Posterior samples: 4000 draws\n")
cat("  Total data transfer:", 6 * 4000 * 8, "bytes (", 
    round(6 * 4000 * 8 / 1024, 1), "KB)\n")
```

#### Path B: Share Only Summary Statistics (requires assumptions)
```{r share_summaries}
# Sites share only mean and SE
# Minimal data transfer, maximum privacy
# BUT: Only valid if posteriors are approximately normal!

summary_only <- site_summaries %>%
  select(site, beta_age_mean, beta_age_se)

cat("Data shared per site:\n")
cat("  Point estimate: 1 number\n")
cat("  Standard error: 1 number\n")
cat("  Total data transfer:", 6 * 2 * 8, "bytes (", 
    round(6 * 2 * 8 / 1024, 3), "KB)\n")

print(summary_only)
```

**Privacy consideration**: Path B shares 190× less data than Path A!   

**Critical assumption**: Path B relies on the **Bayesian Central Limit Theorem**:   
- Large sample sizes at each site   
- Parameters not near boundaries   
- Well-behaved likelihoods   
- Regular posterior geometry   

## Central Coordinator: Stage 2 Shrinkage

The coordinator (e.g., a coordinating center or trusted third party) now applies hierarchical shrinkage.

### Path A: Using Full Posteriors

```{r path_a_mixture}
# Fit mixture approximation
mix <- fit_mixture(
  samples = site_posteriors,
  K_max = 2,  # Age effects should be fairly normal
  verbose = TRUE
)

# Check quality
plot(mix, draws = site_posteriors, type = "density")
```

#### Specify Network-Level Priors

```{r network_priors}
# Based on clinical knowledge:
# - Age effect should be positive but moderate (β₁ ≈ 0.02-0.08)
# - Some heterogeneity expected across hospital types

hierarchical_priors <- list(
  mu = dist_normal(0.05, 0.025),  # Centered on 5% increase per year
  tau = dist_truncated(dist_student_t(3, 0, 0.01), lower = 0)  # Modest heterogeneity
)

# Visualize prior implications
prior_pred <- sample_prior_predictive(
  hierarchical_priors = hierarchical_priors,
  n_groups = 6,
  n_draws = 1000
)

plot(prior_pred, type = "both")
```

#### Fit Hierarchical Model

```{r fit_path_a, results='hide'}
fit_full_post <- shrink(
  mixture = mix,
  hierarchical_priors = hierarchical_priors,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 123
)
```

```{r path_a_results}
print(fit_full_post)

# Network-level estimates
mu_tau_full <- extract_mu_tau(fit_full_post)
cat("\nNetwork-level age effect (μ):\n")
cat("  Posterior mean:", round(mean(mu_tau_full$mu), 4), "\n")
cat("  95% CI: [", round(quantile(mu_tau_full$mu, 0.025), 4), ",",
    round(quantile(mu_tau_full$mu, 0.975), 4), "]\n")

cat("\nBetween-site heterogeneity (τ):\n")
cat("  Posterior mean:", round(mean(mu_tau_full$tau), 4), "\n")
cat("  95% CI: [", round(quantile(mu_tau_full$tau, 0.025), 4), ",",
    round(quantile(mu_tau_full$tau, 0.975), 4), "]\n")
```

### Path B: Using Only Summary Statistics

**Before using summary statistics**, we must verify posteriors are approximately normal!

#### Step 1: Check Normality Assumption

```{r check_normality, fig.width=10, fig.height=8}
# Visual checks for approximate normality
par(mfrow = c(3, 2))
for(i in 1:6) {
  site_name <- names(site_posteriors)[i]
  samples_i <- as.vector(site_posteriors[[i]])
  
  # QQ plot against normal
  qqnorm(samples_i, main = paste("QQ Plot:", site_name))
  qqline(samples_i, col = "red", lwd = 2)
}
par(mfrow = c(1, 1))
```

```{r normality_tests}
# Quantitative checks
normality_checks <- data.frame(
  site = names(site_posteriors),
  skewness = sapply(site_posteriors, function(x) {
    n <- length(x)
    m3 <- mean((x - mean(x))^3)
    s3 <- sd(x)^3
    m3 / s3
  }),
  kurtosis = sapply(site_posteriors, function(x) {
    n <- length(x)
    m4 <- mean((x - mean(x))^4)
    s4 <- sd(x)^4
    m4 / s4 - 3  # Excess kurtosis
  })
)

print(normality_checks)

cat("\nNormality assessment:\n")
cat("  Skewness close to 0? (|skew| < 0.5 is good)\n")
cat("  Kurtosis close to 0? (|kurt| < 1.0 is good)\n")
cat("  All sites pass:", 
    all(abs(normality_checks$skewness) < 0.5 & abs(normality_checks$kurtosis) < 1.0),
    "\n")
```

**Decision rule**:    
- If posteriors look approximately normal → Path B is likely valid   
- If posteriors are skewed/heavy-tailed → Use Path A (mixture)   
- If unsure → Use Path A (safer choice)   

#### When CLT Fails: A Counter-Example

To illustrate why normality matters, consider a scenario where we're estimating a **variance parameter**:

```{r clt_failure_example}
# Simulated: posterior for a variance parameter (boundary at 0)
set.seed(999)
variance_posterior <- matrix(rchisq(4000, df = 5) / 5, ncol = 1)

# Compute summary statistics
var_mean <- mean(variance_posterior)
var_se <- sd(variance_posterior)

# Check normality
par(mfrow = c(1, 2))
hist(variance_posterior, breaks = 30, main = "Variance Posterior",
     xlab = "σ²", col = "lightblue")
qqnorm(variance_posterior, main = "QQ Plot")
qqline(variance_posterior, col = "red", lwd = 2)
par(mfrow = c(1, 1))

# Skewness
skew <- mean((variance_posterior - var_mean)^3) / var_se^3
cat("Skewness:", round(skew, 2), "(should be ~0 for normal)\n")
cat("This posterior is RIGHT-SKEWED - CLT approximation would be poor!\n")
```

**For this scenario:**   
- Path B (summaries) would give biased results   
- Path A (mixture) handles the skewness correctly   

#### Step 2: Fit Using CLT Approximation

```{r path_b_mle}
# Extract means and variances
mle_estimates <- site_summaries$beta_age_mean
names(mle_estimates) <- site_summaries$site

mle_variances <- site_summaries$beta_age_se^2
names(mle_variances) <- site_summaries$site

# Fit using MLE path (CLT approximation)
fit_summaries <- shrink(
  mle = mle_estimates,
  var_matrix = mle_variances,
  hierarchical_priors = hierarchical_priors,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  seed = 123,
  verbose = FALSE,
  refresh = 0
)

print(fit_summaries)
```

### Compare Paths

```{r compare_paths}
mu_tau_summaries <- extract_mu_tau(fit_summaries)

comparison <- data.frame(
  parameter = c("mu", "tau"),
  full_posteriors = c(
    mean(mu_tau_full$mu),
    mean(mu_tau_full$tau)
  ),
  summaries_only = c(
    mean(mu_tau_summaries$mu),
    mean(mu_tau_summaries$tau)
  )
) %>%
  mutate(difference = abs(full_posteriors - summaries_only))

print(comparison)

cat("\nMaximum difference:", round(max(comparison$difference), 5), "\n")
```

**Conclusion**: Both paths give nearly identical results **because** posteriors are approximately normal in this case. This will not always be true!

**When each path is appropriate:**

| Situation | Recommended Path | Reason |
|-----------|------------------|---------|
| Posteriors are normal (verified) | Path B acceptable | CLT holds; minimal sharing |
| Posteriors are skewed/multimodal | Path A required | CLT fails; mixture needed |
| Small sample sizes per site | Path A safer | CLT may not hold yet |
| Boundary constraints (e.g., σ > 0) | Path A required | CLT assumes interior parameters |
| Unknown posterior shape | Path A safer | Conservative choice |
| Maximum privacy needed + normal posteriors | Path B acceptable | But verify normality! |

## Results: Improved Site-Specific Estimates

### Visualize Shrinkage Effect

```{r viz_shrinkage, fig.width=10, fig.height=6}
# Using Path A results (nearly identical for Path B)
plot(fit_full_post)
```

**Key insights:**   
- **Community Hospital** (smallest site) shrinks most toward network mean   
- **University Medical** (largest site) keeps closest to original estimate  
- **Metro General** and **VA Hospital** pulled slightly toward center   
- Adaptive borrowing based on precision!   

### Quantify Uncertainty Reduction

```{r uncertainty_reduction}
# Get Stage 2 estimates
theta_post <- summarize_theta(fit_full_post)

# Compare Stage 1 vs Stage 2 uncertainty
uncertainty_comparison <- data.frame(
  site = site_summaries$site,
  n_patients = site_summaries$n_patients,
  stage1_se = site_summaries$beta_age_se,
  stage2_se = theta_post$sd
) %>%
  mutate(
    reduction_pct = 100 * (stage1_se - stage2_se) / stage1_se,
    stage1_ci_width = 2 * 1.96 * stage1_se,
    stage2_ci_width = 2 * 1.96 * stage2_se,
    ci_width_reduction = 100 * (stage1_ci_width - stage2_ci_width) / stage1_ci_width
  )

print(uncertainty_comparison)
```

**Largest improvements** in smaller sites:   
- Community Hospital: 26% reduction in SE   
- Children's Specialty: 18% reduction in SE   
- County Regional: 15% reduction in SE   

### Visualize Uncertainty Reduction

```{r plot_uncertainty, fig.width=10, fig.height=6}
# Prepare data for plotting
uncertainty_long <- uncertainty_comparison %>%
  select(site, n_patients, stage1_se, stage2_se) %>%
  pivot_longer(
    cols = c(stage1_se, stage2_se),
    names_to = "stage",
    values_to = "standard_error"
  ) %>%
  mutate(
    stage = factor(stage, 
                   levels = c("stage1_se", "stage2_se"),
                   labels = c("Stage 1 (Independent)", "Stage 2 (Shrunken)"))
  )

ggplot(uncertainty_long, aes(x = reorder(site, -n_patients), y = standard_error, 
                              fill = stage)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = sprintf("%.4f", standard_error)),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_manual(values = c("Stage 1 (Independent)" = "steelblue",
                                "Stage 2 (Shrunken)" = "coral")) +
  labs(
    title = "Uncertainty Reduction Through Federated Learning",
    subtitle = "Sites ordered by sample size (largest to smallest)",
    x = "Hospital Site",
    y = "Standard Error (β₁)",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```

## Clinical Impact: Network-Calibrated Predictions

### Stage 1: Independent Site Predictions

```{r stage1_predictions}
# Example: 70-year-old patient
age <- 70
baseline_age <- 60  # Reference age

# Stage 1 predictions (independent)
stage1_log_or <- site_summaries$beta_age_mean * (age - baseline_age)
stage1_or <- exp(stage1_log_or)

stage1_preds <- data.frame(
  site = site_summaries$site,
  log_or = stage1_log_or,
  odds_ratio = stage1_or,
  stage = "Independent"
)
```

### Stage 2: Network-Calibrated Predictions

```{r stage2_predictions}
# Stage 2 predictions (network-calibrated)
stage2_log_or <- theta_post$mean * (age - baseline_age)
stage2_or <- exp(stage2_log_or)

stage2_preds <- data.frame(
  site = theta_post$group,
  log_or = stage2_log_or,
  odds_ratio = stage2_or,
  stage = "Network-Calibrated"
)

# Combine
all_preds <- rbind(stage1_preds, stage2_preds)
```

### Visualize Prediction Changes

```{r plot_predictions, fig.width=10, fig.height=6}
ggplot(all_preds, aes(x = site, y = odds_ratio, fill = stage)) +
  geom_col(position = "dodge") +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray30") +
  geom_text(aes(label = sprintf("%.2f", odds_ratio)),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_manual(values = c("Independent" = "steelblue",
                                "Network-Calibrated" = "coral")) +
  labs(
    title = "Predicted Odds Ratio for 70 vs 60 Year-Old Patient",
    subtitle = "Network calibration stabilizes predictions across sites",
    x = "Hospital Site",
    y = "Odds Ratio",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```

**Clinical interpretation**:   
- Independent estimates vary widely (1.42 to 1.81)   
- Network-calibrated estimates are more consistent (1.52 to 1.60)   
- Small sites benefit most from network information   
- Preserves meaningful site-specific variation   

## Privacy-Preserving Benefits

### What Gets Shared

```{r privacy_summary}
privacy_comparison <- data.frame(
  approach = c("Centralized Data", "Path A: Full Posteriors", "Path B: Summaries"),
  patient_data_shared = c("Yes - All records", "No", "No"),
  data_per_site = c("~50-200 MB", "~200 KB", "~16 bytes"),
  privacy_risk = c("High", "Low", "Minimal"),
  validity = c("N/A", "Always valid", "Only if CLT holds"),
  when_to_use = c("Not for federated", "Default choice", "When posteriors normal")
)

knitr::kable(privacy_comparison, align = "lccccl")
```

**Key insight**: Path A (full posteriors at ~200 KB) is still 1000× smaller than centralized data (~200 MB) while being valid in all scenarios. Path B saves another 12,500× but requires normality.

### Compliance Benefits

**Certain data privacy policies satisfied**:   
- No patient-level data leaves sites     
- Only aggregate statistical summaries shared   
- De-identified parameter estimates   

## Advanced Federated Scenarios

### Scenario 1: Heterogeneous Models

Sites use different model specifications:

```{r heterogeneous_models, eval=FALSE}
# Site 1: Linear model
# Site 2: GLM with splines
# Site 3: Bayesian hierarchical model

# As long as they all estimate the same parameter (e.g., treatment effect),
# shrinkr can combine them!

samples_heterogeneous <- list(
  site1 = samples_from_lm,
  site2 = samples_from_glm,  # Extract coefficient for treatment
  site3 = samples_from_bayes  # Extract theta[treatment]
)

# Proceed with shrinkr as usual
mix <- fit_mixture(samples_heterogeneous, K_max = 3)
fit <- shrink(mix, hierarchical_priors = priors)
```

### Scenario 2: Meta-Analysis of Published Studies

Combine published results without raw data:

```{r meta_analysis, eval=FALSE}
# Extracted from publications
published_estimates <- c(
  "Smith et al. (2020)" = 0.45,
  "Jones et al. (2021)" = 0.52,
  "Garcia et al. (2022)" = 0.38,
  "Williams et al. (2023)" = 0.48
)

published_ses <- c(0.12, 0.15, 0.10, 0.13)

# Apply shrinkr for Bayesian meta-analysis
# NOTE: This assumes published estimates are approximately normal!
# This is usually reasonable for:
#   - Large studies (CLT has kicked in)
#   - Parameters on unrestricted scale (e.g., log-OR, not OR)
#   - Standard errors are from asymptotic approximations

fit_meta <- shrink(
  mle = published_estimates,
  var_matrix = published_ses^2,
  hierarchical_priors = priors
)

# CAUTION: If studies report:
#   - Odds ratios (not log-OR) - use log-scale
#   - Proportions near 0 or 1 - may be skewed
#   - Small sample sizes - CLT may not hold
# In these cases, request posterior samples if possible
```

### Scenario 3: Iterative Federated Updates

New sites join the network over time:

```{r iterative_updates, eval=FALSE}
# Initial network (3 sites)
fit_initial <- shrink(samples_initial, priors)

# New site joins - refit with 4 sites
samples_updated <- c(samples_initial, list(new_site = new_samples))
fit_updated <- shrink(samples_updated, priors)

# Compare network estimates before/after
mu_before <- mean(extract_mu_tau(fit_initial)$mu)
mu_after <- mean(extract_mu_tau(fit_updated)$mu)
```

## Federated Learning Best Practices

### 1. Establish Data Governance

**Before sharing any summaries:**   
- Define what parameters will be shared   
- Establish data use agreements for summary statistics   
- Document privacy protections   
- Get institutional approval   

### 2. Standardize Stage 1 Models

**Ensure comparability:**   
- Use consistent outcome definitions   
- Standardize predictor coding/scaling     
- Document any site-specific modifications   
- Agree on parameter naming conventions   

### 3. Verify Normality (if using summaries

**If using Path B (summary statistics only), sites must:**
- Generate QQ plots of posteriors   
- Compute skewness and excess kurtosis   
- Share diagnostic plots with coordinator   
- Agree normality is reasonable before proceeding

**Red flags for non-normality:**   
- Boundary constraints (variance parameters, probabilities)   
- Small sample sizes (n < 30-50 per site)   
- Highly skewed or heavy-tailed data   

**When in doubt, use Path A** (mixture approximation).

### 4. Quality Control

**Central coordinator should:**   
- Check for outliers in shared summaries   
- Verify mixture approximation quality (Path A)   
- Assess prior-data conflicts   
- Report back site-specific diagnostics   

```{r quality_control}
# Example: Flag suspicious estimates
qc_results <- site_summaries %>%
  mutate(
    z_score = (beta_age_mean - median(beta_age_mean)) / mad(beta_age_mean),
    flag = ifelse(abs(z_score) > 3, "Review", "OK")
  )

cat("Quality control flags:\n")
print(qc_results %>% select(site, beta_age_mean, z_score, flag))
```

### 5. Sensitivity Analysis

Test robustness to prior specifications:

```{r sensitivity, results='hide'}
# Alternative prior: More heterogeneity
priors_alt <- list(
  mu = dist_normal(0.05, 0.025),
  tau = dist_truncated(dist_student_t(3, 0, 0.02), lower = 0)  # Wider tau
)

fit_alt <- shrink(
  mixture = mix,
  hierarchical_priors = priors_alt,
  chains = 2,
  iter = 1000,
  warmup = 500,
  seed = 456
)
```

```{r sensitivity_compare}
# Compare key results
mu_base <- mean(extract_mu_tau(fit_full_post)$mu)
mu_alt <- mean(extract_mu_tau(fit_alt)$mu)

tau_base <- mean(extract_mu_tau(fit_full_post)$tau)
tau_alt <- mean(extract_mu_tau(fit_alt)$tau)

cat("Sensitivity to prior on tau:\n")
cat("  mu: Base =", round(mu_base, 4), ", Alternative =", round(mu_alt, 4), "\n")
cat("  tau: Base =", round(tau_base, 4), ", Alternative =", round(tau_alt, 4), "\n")
```

### 6. Transparent Reporting

**Share with network participants:**   
- Network-level estimates (μ, τ)   
- Site-specific shrunken estimates   
- Diagnostics (convergence, sensitivity)   
- Quantified uncertainty reduction   

```{r reporting}
# Create site-specific report
site_report <- data.frame(
  site = theta_post$group,
  original_estimate = site_summaries$beta_age_mean,
  original_se = site_summaries$beta_age_se,
  calibrated_estimate = theta_post$mean,
  calibrated_se = theta_post$sd,
  uncertainty_reduction = uncertainty_comparison$reduction_pct
) %>%
  mutate(across(where(is.numeric), ~round(.x, 4)))

cat("\nFederated Learning Results Report\n")
cat("==================================\n\n")
cat("Network-level estimate (μ):", round(mean(mu_tau_full$mu), 4), "\n")
cat("Between-site heterogeneity (τ):", round(mean(mu_tau_full$tau), 4), "\n\n")
cat("Site-specific calibrated estimates:\n")
print(site_report)
```



## Advantages of shrinkr for Federated Learning

| Feature | Benefit |
|---------|---------|
| **Two-stage design** | Clean separation between local (Stage 1) and collaborative (Stage 2) analysis |
| **Flexible sharing options** | Can share full posteriors, mixture approximations, OR summaries (if CLT holds) |
| **Privacy preserving** | No patient-level data exposure |
| **Flexible Stage 1** | Each site can use their preferred modeling approach |
| **Transparent shrinkage** | Sites understand how their estimates are adjusted |
| **Uncertainty quantification** | Proper propagation of both within-site and between-site uncertainty |
| **Handles non-normality** | Mixture approximation works for skewed/multimodal posteriors |
| **Regulatory friendly** | Complies with HIPAA, GDPR, and institutional policies |

## When to Use Federated shrinkr

**Ideal scenarios:**   
- Multi-center clinical trials   
- Hospital network collaborations   
- International consortia   
- Meta-analyses with limited published data   
- Any setting where data cannot be centralized

**Requirements:**
- Sites can fit Bayesian models independently   
- Sites can share posterior samples OR mixture approximations   
- **OR** sites can share means + SEs **AND** posteriors are approximately normal   
- Common parameter of interest across sites   
- Central coordinator to run Stage 2   

**Not recommended when:**   
- Sites have very different populations (pooling may not make sense)   
- Sites cannot agree on parameter definitions   
- Posteriors are highly non-normal AND sites cannot share full posteriors/mixtures

## Summary

The shrinkr package enables **privacy-preserving federated learning** through its two-stage design:   

1. **Stage 1** (at each site): Fit local models behind firewall   
2. **Check normality** (if using summaries): Verify CLT assumptions hold   
3. **Share**: Full posteriors (Path A, ~200KB) or summaries (Path B, ~16 bytes)    
4. **Stage 2** (central): Apply hierarchical shrinkage   
5. **Return**: Improved site-specific estimates   

**Key advantages:**   
- Data sovereignty preserved   
- Flexible sharing (full posteriors OR summaries)   
- Handles non-normal posteriors (via mixture approximation)   
- Improved estimates for all sites   
- Proper uncertainty quantification   
- Compliant with privacy regulations   

**Critical decision: Which path?**   
- **Path A** (posteriors/mixture): Always valid, handles any posterior shape   
- **Path B** (summaries): Valid only if posteriors are approximately normal   
- **When uncertain**: Use Path A (safer, still minimal data sharing)   


## Additional Resources

```{r resources, eval=FALSE}
# See also:
vignette("getting_started")  # Core shrinkr workflow
vignette("tidy_bayesian_workflow")  # Working with posteriors
vignette("brms_integration")  # Using brms for Stage 1
```

## Session Info

```{r sessioninfo}
sessionInfo()
```
