---
title: "Meta-Analytic-Predictive (MAP) Priors with shrinkr and beastt"
author: "Jacob M. Maronge"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Meta-Analytic-Predictive (MAP) Priors with shrinkr and beastt}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  message    = FALSE,
  warning    = FALSE,
  fig.width  = 7,
  fig.height = 4.5,
  dpi        = 96
)
```

## Overview

A **meta-analytic-predictive (MAP) prior** summarizes several historical
control arms into a single prior for the control mean of a new trial
(Neuenschwander et al., 2010). The recipe is a random-effects meta-analysis:
study-specific control means are treated as exchangeable draws from a common
population, and the prior for the next trial is the posterior predictive
distribution for a new, as-yet-unobserved study. To protect against prior-data
conflict, the MAP is then robustified by mixing in a vague component
(Schmidli et al., 2014), so that the historical data are automatically
down-weighted when they disagree with the current trial.

This vignette shows how `shrinkr` and `beastt` divide that work cleanly:

- **`shrinkr`** runs the hierarchical meta-analysis across historical studies
  (its Stage-2 shrinkage step) and returns the MAP as a `distributional`
  object.
- **`beastt`** turns that MAP into a robust mixture prior
  (`robustify_norm()`), combines it with the internal control arm to form the
  posterior (`calc_post_norm()`), and computes the effective sample size (ESS).

The two packages meet at a single, intentionally simple hand-off: a `dist_normal`
object. Everything upstream of it is hierarchical modeling (`shrinkr`);
everything downstream is borrowing and inference for the current trial
(`beastt`).

We focus on a continuous outcome with **known** within-arm standard deviation,
which keeps every step conjugate and lets us see the moving parts without
Stan getting in the way.

```{r libraries}
library(shrinkr)
library(beastt)
library(distributional)
library(dplyr)
library(tibble)
library(ggplot2)

set.seed(1104)
```

## Setting

Imagine we are designing a new trial and want to borrow from six historical
studies that each reported a control arm. We have a current ("internal")
control arm of 70 patients, and we assume a known response SD
(`sigma = 2`) in every arm.

```{r truth}
sigma_known <- 2      # known within-arm response SD (same in all arms)
mu_grand    <- 10     # true population-level control mean
tau_true    <- 1.2    # true between-study SD (genuine heterogeneity)
```

The historical studies are *not* identical: each has its own underlying
control mean drawn around the grand mean, which is precisely the
between-study heterogeneity a MAP prior is meant to capture.

```{r simulate-historical}
hist_specs <- tibble(
  study = paste0("H", 1:6),
  n     = c(55, 40, 65, 30, 50, 45),
  # study-specific true means ~ N(mu_grand, tau_true^2)
  theta = rnorm(6, mu_grand, tau_true)
)

# Simulate patient-level control data for each historical study, then reduce
# each study to the summary a publication would actually report: a mean and
# its standard error. (With known sigma, se = sigma / sqrt(n).)
hist_summ <- hist_specs %>%
  rowwise() %>%
  mutate(
    ybar = mean(rnorm(n, theta, sigma_known)),
    se   = sigma_known / sqrt(n)
  ) %>%
  ungroup() %>%
  select(study, n, ybar, se)

hist_summ
```

```{r simulate-internal}
# The current trial's control arm.
n_int   <- 70
int_ctrl <- tibble(
  subjid = seq_len(n_int),
  y      = rnorm(n_int, mu_grand, sigma_known)
)
```

A quick forest-style look at the historical evidence we are about to pool:

```{r forest, fig.height = 3.6}
ggplot(hist_summ, aes(x = study, y = ybar)) +
  geom_hline(yintercept = mean(hist_summ$ybar), linetype = "dashed",
             color = "grey50") +
  geom_pointrange(aes(ymin = ybar - 1.96 * se, ymax = ybar + 1.96 * se),
                  color = "steelblue", linewidth = 0.8) +
  labs(x = NULL, y = "Control mean",
       title = "Historical control arms",
       caption = "Points: study means; bars: 95% intervals; dashed: simple average") +
  theme_minimal(base_size = 12)
```

## Stage 1: per-study estimates

`shrinkr` uses a two-stage workflow. **Stage 1** produces a per-group estimate
of the quantity to be shrunk — here, each study's control mean — using a flat
prior, so that all of the shrinkage happens in Stage 2. With a known SD and a
flat prior on the mean, each study's Stage-1 posterior is exactly
$N(\bar y_g,\, se_g^2)$, so the published summaries `(ybar, se)` *are*
the Stage-1 result. No model fitting is required, and we can pass them straight
to `shrink()` through its summary-statistics interface.

## Stage 2: hierarchical meta-analysis with `shrinkr`

The hierarchical model is

$$
\hat\theta_g \mid \theta_g \sim N(\theta_g,\, se_g^2), \qquad
\theta_g \mid \mu, \tau \sim N(\mu, \tau^2),
$$

with priors on the population mean `mu` and the between-study SD `tau`. We use
a vague prior on `mu` and a weakly-informative half-normal on `tau`.

```{r priors}
hierarchical_priors <- list(
  mu  = dist_normal(0, 100),                                  # vague location
  tau = dist_truncated(dist_normal(0, sigma_known / 2),       # half-normal
                       lower = 0)
)
```

### Check what the heterogeneity prior implies

Before fitting, it is worth seeing what the prior on `tau` implies about
differences *between* study means. `sample_prior_predictive()` draws
`(mu, tau)` and then study effects `theta_g = mu + tau * z_g`;
`prior_pairwise_differences()` summarizes the implied $|\theta_i - \theta_j|$.
Because the location `mu` cancels in any difference, these diagnostics isolate
the heterogeneity prior even though `mu` is vague.

```{r prior-predictive}
prior_pred <- sample_prior_predictive(
  hierarchical_priors = hierarchical_priors,
  n_groups = nrow(hist_summ),
  n_draws  = 2000
)

pw <- prior_pairwise_differences(prior_pred)
print(pw)
plot(pw)
```

The pairwise summary lets us sanity-check the prior on a clinical scale: are
differences of a few units between historical control means plausible but
large differences unlikely? If the implied spread looks unreasonable, adjust
the `tau` prior here — not after seeing the results.

### Fit

`shrinkr`'s `shrink()` accepts study summaries directly via `mle` (the point
estimates) and `var_matrix` (a vector of per-study variances, treated as
independent).

```{r fit-shrink, results = "hide"}
fit_map <- shrink(
  mle                 = hist_summ$ybar,
  var_matrix          = hist_summ$se^2,
  hierarchical_priors = hierarchical_priors,
  chains  = 4, iter = 4000, warmup = 1000,
  seed    = 2026, refresh = 0, verbose = FALSE
)
```

```{r fit-summary}
summarize_mu_tau(fit_map)
```

`mu` is the borrowed estimate of the population control mean and `tau` is the
estimated between-study heterogeneity.

## Constructing the MAP prior

The MAP prior is the posterior **predictive** distribution for a new study's
control mean. Marginalizing over the posterior of `(mu, tau)`, a Normal
approximation has

$$
\text{mean} = \mathbb{E}[\mu \mid \text{data}], \qquad
\text{variance} = \mathrm{Var}(\mu \mid \text{data}) + \mathbb{E}[\tau^2 \mid \text{data}].
$$

`extract_mu_tau()` returns posterior draws of `mu`, `tau`, and `tau_squared`,
so the MAP is a one-liner.

```{r make-map}
make_map <- function(fit) {
  d <- extract_mu_tau(fit)
  dist_normal(mean(d$mu), sqrt(stats::var(d$mu) + mean(d$tau_squared)))
}

map_prior <- make_map(fit_map)
map_prior
```

The MAP's variance has two pieces: our remaining uncertainty about the central
location (`Var(mu)`) *plus* the predictive spread from genuine heterogeneity
(`E[tau^2]`). That second term is what keeps a MAP honestly wider than a naive
pooled mean.

It is useful to express the MAP's informativeness as a prior **effective
sample size** — for a Normal prior on a mean with known `sigma`, that is
`sigma^2 / Var(prior)`, i.e., how many control patients the prior is "worth".

```{r map-ess}
prior_ess <- function(prior, sigma) sigma^2 / variance(prior)

cat(sprintf("MAP prior: mean = %.2f, sd = %.2f, prior ESS = %.1f patients\n",
            mean(map_prior), sqrt(variance(map_prior)),
            prior_ess(map_prior, sigma_known)))
```

## Robustifying with `beastt`

A MAP prior assumes the new trial is exchangeable with the historical ones. If
that turns out to be wrong, an un-robustified MAP can bias the analysis. The
standard fix (Schmidli et al., 2014) is a **robust mixture prior**: mix the MAP
("informative" component) with a vague component, so the data can overrule the
prior when they conflict.

`beastt::robustify_norm()` does this. Its `n` argument scales the vague
component's variance; passing the MAP's prior ESS makes the vague component
worth a single patient's information (a "unit-information" prior), which is the
conventional choice. We place equal 0.5/0.5 weight on the two components.

```{r robustify}
map_ess <- prior_ess(map_prior, sigma_known)
rmp <- robustify_norm(map_prior, n = map_ess, weights = c(0.5, 0.5))

plot_dist(
  "MAP (informative)" = map_prior,
  "Vague component"   = dist_normal(mix_means(rmp)[["vague"]],
                                    mix_sigmas(rmp)[["vague"]]),
  "Robust mixture"    = rmp
)
```

The robust mixture sits on top of the MAP but carries heavier tails, courtesy
of the vague component — that is the insurance against prior-data conflict.

## The internal control posterior

Now we combine the robust mixture prior with the current trial's control arm
using `calc_post_norm()`. With a known SD, the posterior is again a mixture of
normals, and `beastt` updates the mixture weights for us — automatically
down-weighting the informative component if the internal data disagree with it.

```{r posterior}
post_borrow <- calc_post_norm(
  internal_data = int_ctrl,
  response      = y,
  prior         = rmp,
  internal_sd   = sigma_known
)

# A no-borrowing reference: use only the vague component as the prior.
vague_prior <- dist_normal(mix_means(rmp)[["vague"]], mix_sigmas(rmp)[["vague"]])
post_nobrrw <- calc_post_norm(
  internal_data = int_ctrl,
  response      = y,
  prior         = vague_prior,
  internal_sd   = sigma_known
)

plot_dist(
  "No borrowing"          = post_nobrrw,
  "Borrowing (robust MAP)" = post_borrow
)
```

How much did borrowing buy us? The effective sample size compares the posterior
variance with and without borrowing (Pennello & Thompson, 2008): if the
no-borrowing posterior from `n_int` patients has variance `V0`, a borrowed
posterior with variance `Vb` is as informative as `n_int * V0 / Vb` patients.

```{r ess-table}
ess_post <- n_int * variance(post_nobrrw) / variance(post_borrow)

tibble(
  quantity = c("Posterior mean", "Posterior SD", "Effective sample size"),
  `No borrowing` = c(mean(post_nobrrw), sqrt(variance(post_nobrrw)), n_int),
  `Robust MAP`   = c(mean(post_borrow), sqrt(variance(post_borrow)), ess_post)
) %>%
  mutate(across(where(is.numeric), ~round(., 3)))
```

When the historical and internal data are compatible (as here), the robust MAP
sharpens the control posterior and lifts the effective sample size above the 70
internal controls. Under a prior-data conflict, the mixture weight on the
informative component would drop and the ESS gain would shrink toward zero —
exactly the self-correcting behavior we want.

## Summary

- Use **`shrinkr`** to run the hierarchical meta-analysis across historical
  studies and build the MAP as a `dist_normal` predictive distribution, passing
  the study summaries through the `mle` / `var_matrix` interface.
- Use **`beastt`** to `robustify_norm()` the MAP, form the internal control
  posterior with `calc_post_norm()`, and report the effective sample size.
- Always interrogate the heterogeneity prior with
  `sample_prior_predictive()` / `prior_pairwise_differences()` before fitting,
  and lean on the robust mixture so the data can overrule the historical
  evidence when they disagree.

## References

Neuenschwander, B., Capkun-Niggli, G., Branson, M., & Spiegelhalter, D. J.
(2010). Summarizing historical information on controls in clinical trials.
*Clinical Trials*, 7(1), 5–18.

Schmidli, H., Gsteiger, S., Roychoudhury, S., O'Hagan, A., Spiegelhalter, D., &
Neuenschwander, B. (2014). Robust meta-analytic-predictive priors in clinical
trials with historical control information. *Biometrics*, 70(4), 1023–1032.

Pennello, G., & Thompson, L. (2008). Experience with reviewing Bayesian medical
device trials. *Journal of Biopharmaceutical Statistics*, 18(1), 81–115.

```{r session-info}
sessionInfo()
```
